---
title: "Covid91 vaccine study Final2021"
author: "Isai Bercian"
date: "April 13 2021"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Methods

The first step in isolating the subpopulation.

```{r}
Maledata<- filter(FinalData,sex=="guy")
Maledata$sex<- as.factor(Maledata$sex)
Maledata$infected=factor(Maledata$infected)
Maledata$infected <- relevel(Maledata$infected, "OK")
```

# Males

## Bar Charts

```{r}

Males <- Maledata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The bar chart shows that 340 males who were infected with Covid 19 57% of the males who got Covid 19 were give the drug while 43% of males that were infected got the placebo.

## Numerical Summary

```{r}
table1 <- xtabs(~treatment + infected, data=Maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```

The table shows that about 340 males got Covid 19 out of approximately 10,400 males in the study. The table shows that 3.72% of males who receive the drug got Covid 19 and only 2.55% of the males who got the placebo contracted Covid 19. The table shows that 50% of the males that took the vaccine were more likely to contract Covid 19 this suggests that the vaccine is not effective for males.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi Squared table shows that the p-value of 0.0006181 is so small that it is highly unlikely that the results are by chance.  
$H_0$ The null hypothesis shows that there is no difference if males take the vaccine that it will stop them from contracting Covid 19 or not.  
$H_0$ The alternative hypothesis is that the vaccine does make a difference in a male Whether or not a males contracts Covid 19 or not.Which would reject the null hypothesis.

The odds ratio from the Fisher exact test is 1,47:1 when taking the odds ratio and computing it into a percent of effectiveness by using the formula 0/(1+0)*100 you get 59.5% effectiveness. 

Ultimately the vaccine is 40% effective in the Male population, which is absolutely inefficient.

# Females

```{r}
Femaledata<- filter(FinalData,sex=="gal")
Femaledata$sex<- as.factor(Femaledata$sex)
Femaledata$infected=factor(Femaledata$infected)
Femaledata$infected <- relevel(Femaledata$infected, "OK")
levels(Femaledata$infected)
```

## Bar Charts

```{r}

Females <- Femaledata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(Females,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Females,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The bar chart shows that 607 females who were infected with Covid 19 32% of the females who got Covid 19 were give the drug while 66% of females that were infected got the placebo.


## Numerical Summary

```{r}
table2 <- xtabs(~infected + treatment, data=Femaledata)
rowPerc(table2)
colPerc(table2)
print(table2)
```

The table shows that about 607 females got Covid 19 out of approximately 9,200 females in the study. The table shows that 4.44% of females who receive the drug got Covid 19 and only 8.67% of the females who got the placebo contracted Covid 19. The table shows that 67% of the females that took the vaccine were less likely to contract Covid 19 this suggests that the vaccine is effective for females.


## Inferential Results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

The Chi Squared table shows that the p-value of 0.00000000000000002 is so small that it is highly unlikely that the results are by chance.  
$H_0$ The null hypothesis shows that there is no difference if females take the vaccine that it will stop them from contracting Covid 19 or not.  
$H_0$ The alternative hypothesis is that the vaccine does make a difference in a female Whether or not a females contracts Covid 19 or not. Which would reject the null hypothesis.

The odds ratio from the Fisher exact test is 2.05:1 when taking the odds ratio and computing it into a percent of effectiveness by using the formula 0/(1+0)*100 you get 67% effectiveness. 

Ultimately the vaccine is 67% effective in the Female population, which is efficient.

The vaccine is 67% effective in females.




# LGBTQ


```{r}
LGBTQdata<- filter(FinalData,LGBTQ=="gay")
LGBTQdata$LGBTQ<- as.factor(LGBTQdata$LGBTQ)
LGBTQdata$infected=factor(LGBTQdata$infected)
LGBTQdata$infected <- relevel(LGBTQdata$infected, "OK")
```

## Bar Charts

```{r}

LGBTQ <- LGBTQdata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(LGBTQ,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(LGBTQ,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The bar chart shows that 116 LGBTQ who were infected with Covid 19 71% of the LGBTQ who got Covid 19 were give the drug while 29% of LGBTQ that were infected got the placebo.

## Numerical Summary

```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQdata)
rowPerc(table3)
colPerc(table3)
print(table3)
```

The table shows that about 116 LGBTQ got Covid 19 out of approximately 2,040 LGBTQ in the study. The table shows that 71% of LGBTQ who receive the drug got Covid 19 and only 29% of the LGBTQ who got the placebo contracted Covid 19. The table shows that 29% of the LGBTQ that took the vaccine were less likely to contract Covid 19 this suggests that the vaccine is effective for LGBTQ.

## Inferential Results

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

The Chi Squared table shows that the p-value of 0.0000057 is so small that it is highly unlikely that the results are by chance.  
$H_0$ The null hypothesis shows that there is no difference if LGBTQ take the vaccine that it will stop them from contracting Covid 19 or not.  
$H_0$ The alternative hypothesis is that the vaccine does make a difference in a LGBTQ Whether or not a LGBTQ contracts Covid 19 or not. Which would reject the null hypothesis.

The odds ratio from the Fisher exact test is 0.40:1 when taking the odds ratio and computing it into a percent of effectiveness by using the formula 0/(1+0)*100 you get 29% effectiveness. 

The vaccine is not effective in the LGBTQ population because it is only 29% effective. 

# Drug Users


```{r}
Druggiedata<- filter(FinalData,DrugUser=="yes")
Druggiedata$DrugUser<- as.factor(Druggiedata$DrugUser)
Druggiedata$infected=factor(Druggiedata$infected)
Druggiedata$infected <- relevel(Druggiedata$infected, "OK")
```

## Bar Charts

```{r}

Druggies <- Druggiedata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(Druggies,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Druggies,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The bar chart shows that 117 drug users who were infected with Covid 19 57% of the drug users who got Covid 19 were give the drug while 43% of drug users that were infected got the placebo.

## Numerical Summary

```{r}
table4 <- xtabs(~infected + treatment, data=Druggiedata)
rowPerc(table4)
colPerc(table4)
print(table4)
```

The table shows that about 117 drug users got Covid 19 out of approximately 2,283 drug users in the study. The table shows that 33.3% of drug users who receive the drug got Covid 19 and only 66.6% of the drug users who got the placebo contracted Covid 19. The table shows that 60% of the drug users that took the vaccine were less likely to contract Covid 19 this suggests that the vaccine is effective for drug users.

## Inferential Results

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

The Chi Squared table shows that the p-value of 0.0002273 is so small that it is highly unlikely that the results are by chance.  
$H_0$ The null hypothesis shows that there is no difference if drug users take the vaccine that it will stop them from contracting Covid 19 or not.  
$H_0$ The alternative hypothesis is that the vaccine does make a difference in a drug users Whether or not a drug users contracts Covid 19 or not. Which would reject the null hypothesis.

The odds ratio from the Fisher exact test is 1.83:1 when taking the odds ratio and computing it into a percent of effectiveness by using the formula 0/(1+0)*100 you get 60% effectiveness. 

Ultimately the vaccine is 60% effective in the Drug user population, which is efficient.


The vaccine is 60% effective in the Druggie population.

# Discussion & Conclusion

we found that in the four populations that two populations that had the highest rate of effectiveness. Were the females at 67% and the drug user at 60% while the males population and the LGBTQ population showed that the vaccine was not beneficial to them and were ineffective in protecting them from contracting Covid 19. The males population had a 40% effectiveness rate while the LGBTQ population had a 29% effectiveness rate. Which is not effective.




