---
title: "Covid91 vaccine study Final2021"
author: "Isai Bercian"
date: "April 13 2021"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Methods

The first step in isolating the subpopulation.

```{r}
maledata<- filter(FinalData,sex=="guy")
maledata$sex<- as.factor(maledata$sex)
```

# Males

## Bar Charts

```{r}

Males <- maledata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The bar chart shows that 340 males who were infected with Covid 19 57% of the males who got Covid 19 were give the drug while 43% of males that were infected got the placebo.

## Numerical Summary

```{r}
table1 <- xtabs(~treatment + infected, data=maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```

The table shows that about 340 males got Covid 19 out of approximately 10,400 males in the study. The table shows that 3.72% of males who receive the drug got Covid 19 and only 2.55% of the males who got the placebo contracted Covid 19. The table shows that 50% of the males that took the vaccine were more likely to contract Covid 19 this suggests that the vaccine is not effective for males.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi Squared table shows that the p-value of 0.0006181 is so small that it is highly unlikely that the results are by chance.  
$H_0$ The null hypothesis shows that there is no difference if males take the vaccine that it will stop them from contracting Covid 19 or not.  
$H_0$ The alternative hypothesis is that the vaccine does make a difference in a male Whether or not a males contracts Covid 19 or not.Which would reject the null hypothesis.

The odds ratio from the Fisher exact test is 1,47:1 when taking the odds ratio and computing it into a percent of effectiveness by using the formula 0/(1+0)*100 you get 59.5% effectiveness. 

Ultimately the vaccine is 40% effective in the Male population, which is absolutely inefficient.

# Females

```{r}
femaledata<- filter(FinalData,sex=="gal")
femaledata$sex<- as.factor(femaledata$sex)
```

## Bar Charts

```{r}

Females <- femaledata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(Females,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Females,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

## Numerical Summary

```{r}
table1 <- xtabs(~treatment + infected, data=femaledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```



# LGBTQ



# Drug Users



# Discussion & Conclusion






